Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-05-21 paper_jasa.tex
working on intro still

But the strength of the patterns revealed by plots or individual differences among the audience members is something difficult to quantify. \citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics. This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis.

@@

\begin{figure}[hbt]
%\begin{figurehere}
   \centering
       \scalebox{0.95}{\includegraphics{test_category.pdf}}
       \caption{Lineup plot ($m=20$) for testing $H_0: \beta_k=0$. When the alternative hypothesis is true the plot of the observed data should show a vertical displacement between box plots. Which plot has the biggest vertical displacement between the boxplots?}
       \label{fig:test_category}
\end{figure}
%\end{figurehere}

>>>>

However, measuring the strength of patterns seen in plots, and differences in individual perceptual ability, is something that is difficult and perhaps handicaps graphics  use among statisticians, where measuring probabilities is of primary importance. This has also been addressed in recent research. \citet{buja:2009} proposes a protocol that allow the testing of discoveries made from statistical graphics. This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis. One of the protocols, the lineup, places the data plot among  a page of plots of null data, and asks a human judge to pick the plot that is different. Figure \ref{fig:test_category} shows an example lineup. Which plot do you think is the most different from the others? Wrapped in a process that mirrors classical inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis. The null hypothesis typically derives from the task at hand, or the type of plot being made. The alternative encompasses all possible antitheses, all types of patterns that might be detected in the data plot, accounting for all possible deviations from the null without the requirement to specify these ahead of time. The probability of rejection can be quantified, along with Type I, and Type II error, and power can be defined and estimated.

\begin{figure}[htp]
%\begin{figurehere}
   \centering
       \includegraphics[width=0.95\textwidth]{test_category.pdf}
       \caption{Lineup plot ($m=20$) for testing $H_0: \beta_k=0$. When the alternative hypothesis is true the plot of the observed data should show a vertical displacement between box plots. Which plot has the biggest vertical displacement between the boxplots?}
       \label{fig:test_category}
\end{figure}
%\end{figurehere}

The protocol has only been informally tested until now. This paper describes research where the lineup protocol is compared head to head with the relevant classical test. Specifically, the lineup is examined in the context of a linear model setting, where we are determining the importance of including a variable in the model. This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario. The intended use of lineups is where there is no existing test, and unlikely to be ever any numerical test. The thought is though, that the classical setting provides a benchmark for how well the lineup protocol works.

@@
